---
title: "Geospatial data dashboard"
output: html_notebook
---

Load packages.

```{r}
is.installed <- function(paquete) is.element(paquete, installed.packages())

if(!is.installed('rgdal'))

  install.packages('rgdal')

library(rgdal)

if(!is.installed('tmap'))

  install.packages('tmap')

library(tmap)
```

Load CSV data

```{r}
customers <- read.csv2("data/LondonCustomer.csv")
str(customers)
```

Add total product consumption and store it into a variable
```{r}
customers$SUM = customers$NETPRICE_PRO11_AMT + customers$NETPRICE_PRO12_AMT + customers$NETPRICE_PRO13_AMT + customers$NETPRICE_PRO14_AMT + customers$NETPRICE_PRO15_AMT + customers$NETPRICE_PRO16_AMT + customers$NETPRICE_PRO17_AMT

customers$LT_55 = customers$AGE < 55;
str(customers)
customers
```


Select customers that are less than 55 years old and group them by office

```{r}
customer_data = customers[customers$LT_55 == TRUE, ]
customer_data = customer_data[order(customer_data$SUM),]
customer_data
```  

Select branches with lowest grades and group them by region

```{r}
if(!is.installed('dplyr'))

  install.packages('dplyr')
library("dplyr")

customer_data_lowest <- customer_data %>% 
  count(name, SUM, sort = TRUE) %>%
  filter(SUM == 1)

customer_data_lowest <- customer_data_lowest[c(1,3)]
colnames(customer_data_lowest) <- c("name","TotalLowestVolume")
customer_data_lowest



```
Average the total grade per  name
```{r}
customer_data_avg <- customer_data %>% 
  group_by(name) %>% 
  summarise(average_sum = mean(SUM))
customer_data_avg
```

Load geo spatial data

```{r}
i_data_gs <- readOGR(dsn = "data", layer = "london_sport")

i_data_gs$Pop_2001 <- as.numeric(as.character(i_data_gs$Pop_2001))
```

Join geospatial data by region name
```{r}
customer_ag_data <- merge(i_data_gs@data,customer_data_lowest,by="name",all.x=T)
customer_ag_data <- merge(customer_ag_data,customer_data_avg,by="name",all.x=T)
customer_ag_data
```


Data cleanup

```{r}
customer_ag_data$TotalLowestVolume[is.na(customer_ag_data$TotalLowestVolume)] <- 0
customer_ag_data
str(customer_ag_data)
```
Create a new data frame with geospatial data and customer data

```{r}
customer_ag_data_gs<-merge(i_data_gs, customer_ag_data[,c(1,5:6)], by = "name", all.x=TRUE)
customer_ag_data_gs
```

Display customer with lower value per region (using <55 year old filter)

```{r}
qtm(customer_ag_data_gs, "TotalLowestVolume",  text="name", text.size="AREA",text.root=4, fill.title="Customer value", style="gray") +
tm_legend(legend.position = c("left", "top"),
          main.title = "Lowest customer(s) per region",
          main.title.position = "right")
```
 

Display grade per region
```{r}
qtm(customer_ag_data_gs, "average_sum", text="name", text.size="AREA",text.root=3, fill.title="Average grade", style="gray") +
tm_legend(legend.position = c("left", "top"),
          main.title = "Average sum per region",
          main.title.position = "right")

```

